---
title: "R Notebook"
output: html_notebook
---

#Natalie Zarrelli, Spring 2019: Plants, Power and Profits
######Mastery Project: Phylogenetic Trees, comparing via lists and Venn diagrams, and then a third tree project with Trait data attached to one of the trees. Park Trees, Street Trees, and Native Trees (via the Adubon list)


```{r}
library(ape)
library(brranching)
library(picante)
library(maps)
library (dplyr)
```

```{r}
getwd()
setwd("/Users/Natalie/Desktop/botany/0botany Mastery")
```


#park trees
```{r}
parktrees <- c("Acer_rubrum", 
"Acer_platanoides",
"Acer_freemanii", 
"Acer_saccharum", 
"Acer_negundo", 
"Acer_saccharinum", 
"Acer_campestre", 
"Acer_palmatum", 
"Acer_pseudoplatanus", 
"Aesculus_hippocastanum", 
"Aesculus_flava", 
"Aesculus_pavia", 
"Aesculus_parviflora", 
"Amelanchier_canadensis", 
"Amelanchier_NA", 
"Amelanchier_laevis", 
"Amelanchier_arborea", 
"Amelanchier_sanguinea", 
"Carpinus_caroliniana", 
"Carpinus_betulus", 
"Carya_ovata", 
"Carya_cordiformis", 
"Carya_glabra", 
"Celtis_occidentalis", 
"Celtis_sinensis", 
"Cercis_canadensis", 
"Chionanthus_virginicus", 
"Baeolophus_bicolor", 
"Quiscalus_quiscula", 
"Hedera_helix", 
"Asplenium_platyneuron", 
"Buteo_jamaicensis", 
"Mimus_polyglottos", 
"Xylobolus_frustulatus", 
"Passer_domesticus", 
"Oxyura_jamaicensis", 
"Aix_sponsa", 
"Turdus_migratorius", 
"Phlebia_tremellosa", 
"Gavia_immer", 
"Rubus_fruticosus", 
"Sciurus_carolinensis", 
"Cyanocitta_cristata", 
"Ilex_opaca", 
"Phalacrocorax_auritus", 
"Catharus_guttatus", 
"Poronidulus_conchifer", 
"Polystichum_acrostichoides", 
"Bisporella_citrina", 
"Stagmomantis_carolina", 
"Pinus_strobus", 
"Pleurotus_ostreatus", 
"Sturnus_vulgaris", 
"Anas_platyrhynchos", 
"Ardea_herodias", 
"Leucothoe_fontanesiana", 
"Sitta_carolinensis", 
"Biscogniauxia_marginata", 
"Columba_livia", 
"Zonotrichia_albicollis", 
"Haemorhous_mexicanus", 
"Fulica_americana", 
"Lonicera_japonica", 
"Accipiter_cooperi", 
"Cardinalis_cardinalis", 
"Bucephala_albeola", 
"Melanerpes_carolinus", 
"Bubo_virginianus", 
"Amphilogia_gyrosa", 
"Colaptes_auratus", 
"Taxus_baccata", 
"Arum_italicum", 
"Sphyrapicus_varius", 
"Sitta_canadensis", 
"Procyon_lotor", 
"Carex_plantaginea", 
"Trametes_gibbosa", 
"Hymenochaetopsis_olivacea", 
"Anas_clypeata", 
"Candelaria_concolor", 
"Lamium_amplexicaule", 
"Reynoutria_japonica", 
"Ficaria_verna", 
"Entodon_seductrix", 
"Phaeophyscia_adiastola", 
"Lecanora_dispersa", 
"Phragmites_australis", 
"Woodsia_obtusa", 
"Trachemys_scripta", 
"Arctium_minus", 
"Pipilo_erythrophthalmus", 
"Phytolacca_americana", 
"Bryum_argenteum", 
"Dryopteris_carthusiana", 
"Artemisia_vulgaris", 
"Branta_canadensis", 
"Baccharis_halimifolia", 
"Allium_vineale", 
"Euonymus_fortunei", 
"Mycosphaerella_colorata", 
"Alliaria_petiolata", 
"Lasionycteris_noctivagans", 
"Helleborus_orientalis", 
"Crocus_vernus", 
"Cornus_mas", 
"Dolichovespula_maculata", 
"Sanguinaria_canadensis", 
"Lamium_purpureum", 
"Prunus_laurocerasus", 
"Spinus_tristis", 
"Melospiza_melodia", 
"Amphasia_interstitialis", 
"Xylocopa_virginica", 
"Pterostichus_mutus", 
"Narcissus_pseudonarcissus", 
"Lindera_benzoin", 
"Mertensia_virginica", 
"Eranthis_hyemalis", 
"Ardea_alba", 
"Calosoma_scrutator", 
"Apis_mellifera", 
"Junco_hyemalis", 
"Pieris_japonica", 
"Spizella_passerina", 
"Narcissus_NA", 
"Zenaida_macroura", 
"Prunus_serotina", 
"Passerella_iliaca", 
"Scilla_siberica", 
"Galanthus_nivalis", 
"Capsella_bursa-pastoris", 
"Podophyllum_peltatum", 
"Forsythia_NA", 
"Micropterus_salmoides", 
"Poecile_atricapillus", 
"Notemigonus_crysoleucas", 
"Dryobates_pubescens", 
"Vinca_minor", 
"Quercus_rubra", 
"Bjerkandera_adusta", 
"Naphrys_pulex", 
"Jasminum_nudiflorum", 
"Lepomis_macrochirus", 
"Syringa_vulgaris", 
"Rhus_glabra", 
"Anas_rubripes", 
"Cornus_florida", 
"Cornus_racemosa", 
"Cornus_amomum", 
"Cornus_sericea", 
"Cornus_alternifolia", 
"Corylus_colurna", 
"Corylus_avellana", 
"Corylus_americana", 
"Crataegus_phaenopyrum", 
"Crataegus_coccinea", 
"Crataegus_mollis", 
"Crataegus_crus-galli", 
"Crataegus_monogyna", 
"Crataegus_NA", 
"Fagus_grandifolia", 
"Fraxinus_americana", 
"Fraxinus_profunda", 
"Fraxinus_pennsylvanica", 
"Fraxinus_excelsior", 
"Fraxinus_ornus", 
"Ginkgo_biloba", 
"Gleditsia_triacanthos", 
"Ilex_crenata", 
"Ilex_glabra", 
"Ilex_verticillata", 
"Juglans_nigra", 
"Liquidambar_styraciflua", 
"Liriodendron_tulipifera", 
"Maclura_pomifera", 
"Magnolia_soulangeana", 
"Magnolia_tripetala", 
"Magnolia_virginiana", 
"Malus_NA", 
"Malus_prunifolia", 
"Malus_hupehensis", 
"Malus_toringo", 
"Malus_baccata", 
"Morus_alba", 
"Nyssa_sylvatica", 
"Phellodendron_amurense", 
"Pinus_rigida", 
"Pinus_wallichiana", 
"Pinus_nigra", 
"Pinus_resinosa", 
"Platanus_occidentalis", 
"Platanus_hispanica", 
"Prunus_persica", 
"Prunus_virginiana", 
"Prunus_yedoensis", 
"Prunus_NA", 
"Quercus_alba", 
"Quercus_palustris", 
"Quercus_cerris", 
"Quercus_bicolor", 
"Quercus_macrocarpa", 
"Quercus_NA", 
"Quercus_ilicifolia", 
"Quercus_coccinea", 
"Quercus_phellos", 
"Quercus_nigra", 
"Quercus_velutina", 
"Quercus_robur", 
"Robinia_pseudoacacia", 
"Taxodium_distichum", 
"Tilia_americana", 
"Tilia_cordata", 
"Tilia_platyphyllos", 
"Ulmus_americana", 
"Ulmus_NA", 
"Ulmus_glabra", 
"Ulmus_pumila", 
"Ulmus_rubra", 
"Ulmus_parvifolia")

parktrees

parktrees_phylo <- phylomatic(taxa = parktrees, storedtree ="zanne2014", get = 'POST')
plot(parktrees_phylo, no.margin = TRUE)
```



```{r}
setwd("/Users/Natalie/Desktop/botany/0botany Mastery")

plot(parktrees_phylo, no.margin = TRUE, type="cladogram")
newick_park = "parktrees_phylo"
write.tree(parktrees_phylo, file=newick_park)
```

```{r}
setwd("/Users/Natalie/Desktop/botany/botany biodiversity project")
```

```{r}
getwd()
```


#street trees

```{r}
streettrees <- c("Acer platanoides", 
"Acer rubrum", 
"Acer saccharinum", 
"Fagus grandifolia", 
"Fraxinus pennsylvanica", 
"Ginkgo biloba", 
"Gleditsia triacanthos", 
"Gymnocladus dioica", 
"Koelreuteria paniculata", 
"Platanus hybrida", 
"Pyrus calleryana", 
"Quercus bicolor", 
"Quercus imbricaria", 
"Quercus palustris", 
"Quercus phellos", 
"Quercus rubra", 
"Quercus velutina", 
"Robinia pseudoacacia", 
"Styphnolobium japonicum", 
"Syringa reticulata", 
"Tilia americana", 
"Tilia cordata", 
"Tilia tomentosa", 
"Ulmus americana", 
"Ulmus parvifolia", 
"Zelkova serrata")

streettrees

streettrees_phylo <- phylomatic(taxa = streettrees, storedtree ="zanne2014", get = 'POST') 
plot(streettrees_phylo, no.margin = TRUE) 

```


```{r}
setwd("/Users/Natalie/Desktop/botany/0botany Mastery")

plot(streettrees_phylo, no.margin = TRUE, type="cladogram") 
newick_street = "streettrees_phylo" 
write.tree(streettrees_phylo, file=newick_street) 
```

#to use with picante
```{r}
newick_streettrees = "((platanus_hybrida:136.9,(((((quercus_phellos:40.41,Fagus_grandifolia:40.41,(Quercus_imbricaria:2.77794,((Quercus_palustris:0.38487,(Quercus_velutina:0.127739,Quercus_rubra:0.127739):0.25713):1.68414,Quercus_bicolor:2.06901):0.708931):37.6321)Fagaceae:65.526,((Zelkova_serrata:38.1072,(Ulmus_americana:37.7515,Ulmus_parvifolia:37.7515):0.355694):51.1699,Pyrus_calleryana:89.2771)Rosales:16.6589):1.05941,(styphnolobium_japonicum:69.1425,((Gymnocladus_dioica:9.45824,Gleditsia_triacanthos:9.45824):48.124,Robinia_pseudoacacia:57.5823):11.5602)Fabaceae:37.8529):10.5393,((Tilia_americana:8.39273,(Tilia_tomentosa:4.03116,Tilia_cordata:4.03116):4.36157):92.1209,((Acer_platanoides:19.0619,(Acer_saccharinum:1.07717,Acer_rubrum:1.07717):17.9848):27.0354,Koelreuteria_paniculata:46.0973):54.4163):17.0211)Rosidae:1.63077,(Fraxinus_pennsylvanica:21.6934,Syringa_reticulata:21.6934):97.4721)Gunneridae:17.735)Eudicotyledoneae:215.334,Ginkgo_biloba:352.235)Spermatophyta;"
street2 = read.tree(text=newick_streettrees)
plot(street2, type="phylogram", show.node.label=TRUE)
```



#Native Tree species NYC

```{r}

nativetrees <- c("Acer negundo",
"Acer rubrum",
"Acer saccharinum",
"Alnus serrulata",
"Amelanchier arborea",
"Baccharis halimifolia",
"Carpinus caroliniana",
"Carya cordiformis",
"Carya glabra",
"Carya ovata",
"Carya tomentosa",
"Castanea dentata",
"Celtis occidentalis",
"Chamaecyparis thyoides",
"Cornus alternifolia",
"Cornus florida",
"Cornus racemosa",
"Crataegus crus-galli",
"Diospyros virginiana",
"Fraxinus americana",
"Fraxinus pennsylvanica",
"Hamamelis virginiana",
"Juglans nigra",
"Juniperus virginiana",
"Kalmia latifolia",
"Larix laricina",
"Liquidambar styraciflua",
"Liriodendron tulipifera",
"Morus rubra",
"Ostrya virginiana",
"Pinus resinosa",
"Pinus rigida",
"Pinus strobus",
"Platanus occidentalis",
"Populus deltoides",
"Populus tremuloides",
"Prunus pensylvanica",
"Prunus serotina",
"Prunus virginiana",
"Ptelea trifoliata",
"Quercus alba",
"Quercus montana",
"Quercus palustris",
"Quercus prinoides",
"Quercus rubra",
"Rhus copallinum",
"Rhus glabra",
"Rhus typhina",
"Salix discolor",
"Salix nigra",
"Sassafras albidum",
"Staphylea trifolia",
"Thuja occidentalis",
"Tsuga canadensis",
"Ulmus americana",
"Viburnum prunifolium")

nativetrees

nativetrees_phylo <- phylomatic(taxa = streettrees, storedtree ="zanne2014", get = 'POST')
plot(nativetrees_phylo, no.margin = TRUE)



```


```{r}

setwd("/Users/Natalie/Desktop/botany/0botany Mastery")
plot(nativetrees_phylo, no.margin = TRUE, type="cladogram")
newick_native = "nativetrees_phylo"
write.tree(nativetrees_phylo, file=newick_native)


```




#A new way to compare the two lists:

```{r}
install.packages("gplots") 
```


```{r}
#first <- c(1:3, letters[1:6], "foo", "bar")
#second <- c(2:4, letters[5:8], "bar", "asd")

#both <- first[first %in% second] # in both, same as call: intersect(first, second)
#onlyfirst <- first[!first %in% second] # only in 'first', same as: setdiff(first, second)
#onlysecond <- second[!second %in% first] # only in 'second', same as: setdiff(second, first)
#length(both) 
#length(onlyfirst) 
#length(onlysecond) 
```


------

Venn Diagram and list contents: 

```{r}
A = c("Acer negundo",
"Acer rubrum",
"Acer saccharinum",
"Alnus serrulata",
"Amelanchier arborea",
"Baccharis halimifolia",
"Carpinus caroliniana",
"Carya cordiformis",
"Carya glabra",
"Carya ovata",
"Carya tomentosa",
"Castanea dentata",
"Celtis occidentalis",
"Chamaecyparis thyoides",
"Cornus alternifolia",
"Cornus florida",
"Cornus racemosa",
"Crataegus crus-galli",
"Diospyros virginiana",
"Fraxinus americana",
"Fraxinus pennsylvanica",
"Hamamelis virginiana",
"Juglans nigra",
"Juniperus virginiana",
"Kalmia latifolia",
"Larix laricina",
"Liquidambar styraciflua",
"Liriodendron tulipifera",
"Morus rubra",
"Ostrya virginiana",
"Pinus resinosa",
"Pinus rigida",
"Pinus strobus",
"Platanus occidentalis",
"Populus deltoides",
"Populus tremuloides",
"Prunus pensylvanica",
"Prunus serotina",
"Prunus virginiana",
"Ptelea trifoliata",
"Quercus alba",
"Quercus montana",
"Quercus palustris",
"Quercus prinoides",
"Quercus rubra",
"Rhus copallinum",
"Rhus glabra",
"Rhus typhina",
"Salix discolor",
"Salix nigra",
"Sassafras albidum",
"Staphylea trifolia",
"Thuja occidentalis",
"Tsuga canadensis",
"Ulmus americana",
"Viburnum prunifolium")

S = c("Acer platanoides", 
"Acer rubrum", 
"Acer saccharinum", 
"Fagus grandifolia", 
"Fraxinus pennsylvanica", 
"Ginkgo biloba", 
"Gleditsia triacanthos", 
"Gymnocladus dioica", 
"Koelreuteria paniculata", 
"Platanus hybrida", 
"Pyrus calleryana", 
"Quercus bicolor", 
"Quercus imbricaria", 
"Quercus palustris", 
"Quercus phellos", 
"Quercus rubra", 
"Quercus velutina", 
"Robinia pseudoacacia", 
"Styphnolobium japonicum", 
"Syringa reticulata", 
"Tilia americana", 
"Tilia cordata", 
"Tilia tomentosa", 
"Ulmus americana", 
"Ulmus parvifolia", 
"Zelkova serrata")

A %in% S
intersect(A,S)
setdiff(A,S)
setdiff(S,A)

```


```{r}
m <- match(A,S)
A[!is.na(m)]

S[m[!is.na(m)]]

```

```{r}
A[is.na(m)]
S[which(is.na(m))]
```


```{r}
P = c("Acer rubrum", 
"Acer platanoides", 
"Acer freemanii", 
"Acer saccharum", 
"Acer negundo", 
"Acer saccharinum", 
"Acer campestre", 
"Acer palmatum", 
"Acer pseudoplatanus", 
"Aesculus hippocastanum", 
"Aesculus flava", 
"Aesculus pavia", 
"Aesculus parviflora", 
"Amelanchier canadensis", 
"Amelanchier NA", 
"Amelanchier laevis", 
"Amelanchier arborea", 
"Amelanchier sanguinea", 
"Carpinus caroliniana", 
"Carpinus betulus", 
"Carya ovata", 
"Carya cordiformis", 
"Carya glabra", 
"Celtis occidentalis", 
"Celtis sinensis", 
"Cercis canadensis", 
"Chionanthus virginicus", 
"Baeolophus bicolor", 
"Quiscalus quiscula", 
"Hedera helix", 
"Asplenium platyneuron", 
"Buteo jamaicensis", 
"Mimus polyglottos", 
"Xylobolus frustulatus", 
"Passer domesticus", 
"Oxyura jamaicensis", 
"Aix sponsa", 
"Turdus migratorius", 
"Phlebia tremellosa", 
"Gavia immer", 
"Rubus fruticosus", 
"Sciurus carolinensis", 
"Cyanocitta cristata", 
"Ilex opaca", 
"Phalacrocorax auritus", 
"Catharus guttatus", 
"Poronidulus conchifer", 
"Polystichum acrostichoides", 
"Bisporella citrina", 
"Stagmomantis carolina", 
"Pinus strobus", 
"Pleurotus ostreatus", 
"Sturnus vulgaris", 
"Anas platyrhynchos", 
"Ardea herodias", 
"Leucothoe fontanesiana", 
"Sitta carolinensis", 
"Biscogniauxia marginata", 
"Columba livia", 
"Zonotrichia albicollis", 
"Haemorhous mexicanus", 
"Fulica americana", 
"Lonicera japonica", 
"Accipiter cooperi", 
"Cardinalis cardinalis", 
"Bucephala albeola", 
"Melanerpes carolinus", 
"Bubo virginianus", 
"Amphilogia gyrosa", 
"Colaptes auratus", 
"Taxus baccata", 
"Arum italicum", 
"Sphyrapicus varius", 
"Sitta canadensis", 
"Procyon lotor", 
"Carex plantaginea", 
"Trametes gibbosa", 
"Hymenochaetopsis olivacea", 
"Anas clypeata", 
"Candelaria concolor", 
"Lamium amplexicaule", 
"Reynoutria japonica", 
"Ficaria verna", 
"Entodon seductrix", 
"Phaeophyscia adiastola", 
"Lecanora dispersa", 
"Phragmites australis", 
"Woodsia obtusa", 
"Trachemys scripta", 
"Arctium minus", 
"Pipilo erythrophthalmus", 
"Phytolacca americana", 
"Bryum argenteum", 
"Dryopteris carthusiana", 
"Artemisia vulgaris", 
"Branta canadensis", 
"Baccharis halimifolia", 
"Allium vineale", 
"Euonymus fortunei", 
"Mycosphaerella colorata", 
"Alliaria petiolata", 
"Lasionycteris noctivagans", 
"Helleborus orientalis", 
"Crocus vernus", 
"Cornus mas", 
"Dolichovespula maculata", 
"Sanguinaria canadensis", 
"Lamium purpureum", 
"Prunus laurocerasus", 
"Spinus tristis", 
"Melospiza melodia", 
"Amphasia interstitialis", 
"Xylocopa virginica", 
"Pterostichus mutus", 
"Narcissus pseudonarcissus", 
"Lindera benzoin", 
"Mertensia virginica", 
"Eranthis hyemalis", 
"Ardea alba", 
"Calosoma scrutator", 
"Apis mellifera", 
"Junco hyemalis", 
"Pieris japonica", 
"Spizella passerina", 
"Narcissus NA", 
"Zenaida macroura", 
"Prunus serotina", 
"Passerella iliaca", 
"Scilla siberica", 
"Galanthus nivalis", 
"Capsella bursa-pastoris", 
"Podophyllum peltatum", 
"Forsythia NA", 
"Micropterus salmoides", 
"Poecile atricapillus", 
"Notemigonus crysoleucas", 
"Dryobates pubescens", 
"Vinca minor", 
"Quercus rubra", 
"Bjerkandera adusta", 
"Naphrys pulex", 
"Jasminum nudiflorum", 
"Lepomis macrochirus", 
"Syringa vulgaris", 
"Rhus glabra", 
"Anas rubripes", 
"Cornus florida", 
"Cornus racemosa", 
"Cornus amomum", 
"Cornus sericea", 
"Cornus alternifolia", 
"Corylus colurna", 
"Corylus avellana", 
"Corylus americana", 
"Crataegus phaenopyrum", 
"Crataegus coccinea", 
"Crataegus mollis", 
"Crataegus crus-galli", 
"Crataegus monogyna", 
"Crataegus NA", 
"Fagus grandifolia", 
"Fraxinus americana", 
"Fraxinus profunda", 
"Fraxinus pennsylvanica", 
"Fraxinus excelsior", 
"Fraxinus ornus", 
"Ginkgo biloba", 
"Gleditsia triacanthos", 
"Ilex crenata", 
"Ilex glabra", 
"Ilex verticillata", 
"Juglans nigra", 
"Liquidambar styraciflua", 
"Liriodendron tulipifera", 
"Maclura pomifera", 
"Magnolia soulangeana", 
"Magnolia tripetala", 
"Magnolia virginiana", 
"Malus NA", 
"Malus prunifolia", 
"Malus hupehensis", 
"Malus toringo", 
"Malus baccata", 
"Morus alba", 
"Nyssa sylvatica", 
"Phellodendron amurense", 
"Pinus rigida", 
"Pinus wallichiana", 
"Pinus nigra", 
"Pinus resinosa", 
"Platanus occidentalis", 
"Platanus hispanica", 
"Prunus persica", 
"Prunus virginiana", 
"Prunus yedoensis", 
"Prunus NA", 
"Quercus alba", 
"Quercus palustris", 
"Quercus cerris", 
"Quercus bicolor", 
"Quercus macrocarpa", 
"Quercus NA", 
"Quercus ilicifolia", 
"Quercus coccinea", 
"Quercus phellos", 
"Quercus nigra", 
"Quercus velutina", 
"Quercus robur", 
"Robinia pseudoacacia", 
"Taxodium distichum", 
"Tilia americana", 
"Tilia cordata", 
"Tilia platyphyllos", 
"Ulmus americana", 
"Ulmus NA", 
"Ulmus glabra", 
"Ulmus pumila", 
"Ulmus rubra", 
"Ulmus parvifolia")
```

```{r}
mm <- match(P,S)
P[!is.na(m)]

S[m[!is.na(m)]]
```
```{r}
A %in% P
intersect(A,P)

#what is in the native audobon list but not the park
setdiff(A,P)
#what is in the park but not in the native list
setdiff(P,A)
```


```{r}
require("gplots")
venn(list(Native.Species = A, Street.Species = S, Park.Species = P))
```

```{r}
length(intersect(A,S))
length(setdiff(A,S))
length(setdiff(S,A))
```


#Using Picante to look at trait data on a tree - Native Trees traits as attributes obtained from the Audubon list (also as excel sheet in the Courseworks page), originally obtained at this link: https://www.audubon.org/native-plants with the zip code 11025. I copied the list (a bit tedious) into excel and formatted the "attribute" data into separate columns. I then saw that I couldn't use the data in word form as a matrix (which gave me errors, I gave it a try anyway) and so converted each character to a numerical value, like a code. This way I could see the traits as descrete, have/have not values. 

#csv data and make a matrix

```{r}
nativetraitslist <- read.csv("nativetrees - Sheet1.csv", row.names = 1)

nativetraitslist
```


#matrix of the csv data

```{r}
nativetraitslist <- as.matrix(nativetraitslist)
dim(nativetraitslist)
```

#native tree plot for picante - to plot clumped traits on the tree using the trait data I formatted in excel and as a matrix

```{r}
newick_nativetrees = "((platanus_hybrida:136.9,(((((quercus_phellos:40.41,Fagus_grandifolia:40.41,(Quercus_imbricaria:2.77794,((Quercus_palustris:0.38487,(Quercus_velutina:0.127739,Quercus_rubra:0.127739):0.25713):1.68414,Quercus_bicolor:2.06901):0.708931):37.6321)Fagaceae:65.526,((Zelkova_serrata:38.1072,(Ulmus_americana:37.7515,Ulmus_parvifolia:37.7515):0.355694):51.1699,Pyrus_calleryana:89.2771)Rosales:16.6589):1.05941,(styphnolobium_japonicum:69.1425,((Gymnocladus_dioica:9.45824,Gleditsia_triacanthos:9.45824):48.124,Robinia_pseudoacacia:57.5823):11.5602)Fabaceae:37.8529):10.5393,((Tilia_americana:8.39273,(Tilia_tomentosa:4.03116,Tilia_cordata:4.03116):4.36157):92.1209,((Acer_platanoides:19.0619,(Acer_saccharinum:1.07717,Acer_rubrum:1.07717):17.9848):27.0354,Koelreuteria_paniculata:46.0973):54.4163):17.0211)Rosidae:1.63077,(Fraxinus_pennsylvanica:21.6934,Syringa_reticulata:21.6934):97.4721)Gunneridae:17.735)Eudicotyledoneae:215.334,Ginkgo_biloba:352.235)Spermatophyta;"
native2 = read.tree(text=newick_nativetrees)
plot(native2, type="phylogram", show.node.label=TRUE)
```


```{r}
traitsNT=read.csv("nativetrees - Sheet1.csv", row.names = 1)
traitsNT
```


#plotting the different descrete traits (each column is an attribute of the tree, each former word, which is indicated as a column name, is represented by a specific number) I was hoping this was a way to view various attributes the trees may have and compare these attributes to the phylogenetic relationships between the native tree species. 

#I found it interesting that all but one non-Ginko Biloba clade had developed shrub habits. Also, all but one species is pertinent to butterflies/caterpillars, and it was interesting to see how the nut trait varied.

#Unfortunately, I'm unsure why I am limited to seven colors; it's not possibly to see the evergreen or seeds traits, which are visible in the matrix data. I attempted to fix this, but kept coming up with the same issue. 
```{r}
par(mfrow = c(2, 2))
for (i in names(traitsNT)) {
plot(native2, show.tip.label = TRUE, main = i)
tiplabels(pch = 22, col = traitsNT[, i] + 1, bg = traitsNT[, i] + 1, cex = 1.5)
}
```

